An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-06 paper.tex
Conclusions

In this paper, we explore the operation of the IETF to understand who
develops Internet standards, how the participants interact, and what
factors may lead to success.
Section \ref{sec:background} introduces the IETF and the Internet Standards
development process, and reviews the data with which we work.
Section \ref{sec:trends-documents} considers trends in document production,
complexity, and correctness.
item Section \ref{sec:trends-demographics} moves on to consider the people
who are involved in Internet standards development and how the demographics
of the community have shifted over time.
Section \ref{sec:org-dyn} studies the interaction between participants,
explores who has influence and impact in the IETF and how that influence is
evident in their activities, and the dynamics of communications within the
IETF.
Section \ref{sec:language} considers communications within the IETF from
the perspective of language to explore whether there are differences in
linguistic patterns and language use in more influential participants, and
to consider how the use of language reflects the consensus-driven nature of
the standards development process.
Section \ref{sec:success-factors} then considers factors that lead to
success, exploring what makes a document likely to be published as an RFC
and what makes a successful participant.
Finally, we review related work in Section \ref{sec:related} and conclude
in Section \ref{sec:conclusions}.

@@

>>>>

In this paper, we explore the operation of the IETF to understand who
develops Internet standards, how the participants interact, and what
factors may lead to success.
Section \ref{sec:background} introduces the IETF and the Internet Standards
development process, and reviews the data with which we work.
Section \ref{sec:trends-documents} considers trends in document production,
complexity, and correctness.
Section \ref{sec:trends-demographics} moves on to consider the people
who are involved in Internet standards development and how the demographics
of the community have shifted over time.
Section \ref{sec:org-dyn} studies the interaction between participants,
explores who has influence and impact in the IETF and how that influence is
evident in their activities, and the dynamics of communications within the
IETF.
Section \ref{sec:language} considers communications within the IETF from
the perspective of language to explore whether there are differences in
linguistic patterns and language use in more influential participants, and
to consider how the use of language reflects the consensus-driven nature of
the standards development process.
Section \ref{sec:success-factors} then considers factors that lead to
success, exploring what makes a document likely to be published as an RFC
and what makes a successful participant.
Finally, we review related work in Section \ref{sec:related} and conclude
in Section \ref{sec:conclusions}.

@@

We explored the key standardisation activities within the IETF, considering
trends in document production, complexity, and correctness
(\S\ref{sec:trends-documents}); the people involved in Internet standards
development and how the demographics of the community have shifted over
time (\S\ref{sec:trends-demographics}); interaction between participants,
who has influence and impact in the IETF and how that influence is evident
in their activities, and the dynamics of communications within the IETF
(\S\ref{sec:org-dyn}); use of language, to explore whether there are
differences in linguistic patterns and language use in more influential
participants, and to consider how the use of language reflects the
consensus-driven nature of the standards development process
(\S\ref{sec:language}); and factors that lead to success, exploring what
makes a document likely to be published as an RFC and what makes a
successful participant (\S\ref{sec:success-factors}).

In the process, we have revealed a vibrant ecosystem, albeit one with a
number of key challenges:
\begin{enumerate}
  \item The rate of standards development has slowed over time and the IETF
    is showing signs of struggling with managing the complexity of
    standards development.

  \item While authorship is becoming more diverse as the IETF grows, there
    are still significant issues with representation from certain regions,
    including Africa and South America. If the IETF is to become more
    representative, then further effort is required to build more diverse
    engagement, both in terms of geography, organisational affiliation, and
    gender diversity \cite{kaeo:2023:women}.

  \item Despite significant growth over time, a relatively small subset of
    well-connected and highly influential participants remain an important
    driving force in the IETF. The community is growing better connected,
    and has become less fragmented, over time, and is growing in resilience
    to departure of influential participants, but there is more to be done.

On the positive side, we observed that influential people in IETF exhibit
collaborative and community-oriented traits, and also stronger signs of
engagement in discussions. We also observed that as people go on to become
influential participants, they evolve in their communication and are seen
to be more engaging and descriptive in their linguistic style and patterns
of communication.  The IETF has a reputation for being a difficult
community to engage with (``loud men, talking loudly'') that is not
altogether unwarranted \cite{cath:2023:loud,cath:2021:the-technology}, but
there is also evidence that those in positions of influence, and leadership
roles, are broadly cooperative and supportive of newcomers and those with
less experience, within their own cultural framework.

Further research is needed to fully understand the IETF and its role in the
Internet governance ecosystem, but our analysis to date shows a complex and
evolving organisation---with some expected growing pains---that is meeting
the ongoing challenges of developing technical standards for the modern,
global, Internet.
